---
title: "working with weirds"
author: "WeLoveDataScience"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{working with weirds}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,  
                      fig.width=10 ,
                      fig.height=7,
                      out.width='100%',
                      dpi=100)
library(knitr)
library(dplyr)
library(ggplot2)
suppressMessages(library(stranger)) 
library(tidyr)
```



# Anomalies manual selection

For some metrics, litterature recommends some cutpoints (2 for `lof` for instance).

There is a way to derive anomalies from this given cutpoint by filtering a stranger or a singularize object and convert to anomalies the set of retained records. 

Filtering can be done either with usual `[` or `dplyr` `select`.

Assumed `dbscan` package is installed, one can for instance use `lof` weird, use 2 as cutpoint and derive an anomaly object with [`as.anomalies`](./../references/as.anomalies.html).


```{r}
iris %>% select(-Species) %>%
  crazyfy() %>%
  strange(weird="lof") %>%
  singularize() -> s
anom <- filter(s,lof_minPts_5>2) %>% as.anomalies()
```

Anomaly objects are simple vectors with attributes; once can then access the first anomaly and display it with:

```{r}
plot(s,type="neighbours", score="lof_minPts_5",anomaly_id=anom[1])
```

# Using a tuneGrid

`stranger` package is a wrapper around several methods. Problem thus arrives on the _choice_ of the method... We will see later in this vignette how to work with several metrics and in particular the stacking.

But world is really complex and even for one chosen method, leyt say `knn` _weird_, user has to pick some parameters...

Here comes the parameter `tuneGrid` (name borrowed from `caret`) that allow to use one _weird_ with different sets of parameters in the same call.

Let's prepare a matrix of parameters:

```{r}
tg <- expand.grid(k=c(5,10,20),simplify=c("mean","median"),stringsAsFactors=FALSE)
tg
```

We thus have 6 possible combinations

```{r}
(anoms.all <-  iris %>% select(-Species) %>%
  crazyfy() %>% strange(weird="knn",tuneGrid=tg))
```

__NOTES__

* `stranger` naming conventions do not necessary involve all parameters, we though ensure unique names and store parameters values as attributes (metadata) for every metric.
* Currently, only parameters exposed in _weird_ method are allowed in tuning grid. User can still invoke methods several times with different values for other parameters and merge results as seen in next section or use `stranger`.


```{r}
tg = data.frame(k=c(5,5:8))
(anoms <-  iris %>% select(-Species) %>%
  crazyfy() %>% strange(weird="knn",tuneGrid=tg,algorithm=c("cover_tree")))
(meta  <- get_info(anoms))
meta[,"parameters"]
```


# Merging stranger objects

So you have decided to go for and try two different methods. Let say _knn_ and _autoencode_.

First, we can create two object containg assciatdd metrics.

```{r}
data <- iris %>% select(-Species) %>% crazyfy()
m1 <- strange(data, weird="knn")
m2 <- strange(data, weird="autoencode")
```

For convenience and further exploitation, a `merge` method is at your disposal to gather the two corresponding metrics in  a single object:

```{r}
(metrics <- merge(m1,m2))
```

# Invoke several weirds at once (stranger) 

You are not sure yet about the metric you want to use. Or you would like to test many, with plenty of different values for the parameters? `stranger` is for you. This function is similar to `caretList` in `caretEnsemble` package.

With its first parameter _weirdList_ you can supply many methods that will be invoked with their default parameter. 

If you want to use your own values or use a tuneGrid for a given method, you will have to create  _weird_ objects using `weird` function and pass a list of such  _weirds_ object to the parameter _tuneList_.


Following call will for instance:
* fit a knn _weird_ with default values (taken from _weirdList_)
* fit a autoencode _weird_ with default values
* also fit a knn but with _k_ set to 20 (not recommended for iris data having 50 observations)

```{r}
s3=stranger(data,
    weirdList=c("knn","autoencode"),
    tuneList=list(weird(method="knn",k=20)))
```



# Normalize & stack metrics

<to be done>.


